Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2015-03-16 MS_revised.tex
cleanned supp info in ms
rename sup mat in tex file

For analyses, we used the phylogenetic trees from the study of Wahlberg
et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
employed the maximum clade credibility tree (MCC tree) (Fig.~\ref{fig1}) as well
as a random sample of 1000 trees from their BEAST run after burnin
\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
used a burnin of 25 million generations and took a random sample of 1000
trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S9 File)
in order to account for phylogenetic uncertainty when performing
the diversification analyses.

@@

Species richness data for Nymphalidae genera were compiled from several
sources including the specialist-curated lists 
\cite{tolweb2007, lamas2004} and curated lists of the
Global Butterfly Names project \cite{gbnp}.
We assigned the species numbers of genera not included in the phylogeny
to the closest related genus that was included in Wahlberg et al. (2009)
\cite{wahlberg2009} study according to available phylogenetic studies 
\cite{matos2013, brower2010, kodandaramaiah2010, kodandaramaiah2010a,
ortiz2013, desilva2010, freitas2004, pena2006, penz1999, silva2008,
pena2011, pena2010},
taxonomical classification and morphological resemblance when no
phylogenies were available (S13 Dataset).

Hostplant data for Nymphalidae species were compiled from several
sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
\cite{dyer2002, beccaloni2008, janzen2009} and others 
(S14 and S15 Dataset) for a
total of 6586
hostplant records, including 428 Nymphalidae genera and 143 plant
families and 1070 plant genera. It was not possible to find any
hostplant data for 35 butterfly genera.

@@

We used the statistical software R version 3.0.1 \cite{r2013} in combination
with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA
\cite{alfaro2009} and
\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
perform the analyses (included as S10 and S11 File). All
analyses were run on the 1000 random trees from Wahlberg et al. (2009)
\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
the MCC tree \cite{wahlberg2009}.

@@

Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
version 093.4.33 \cite{alfaro2009} on the maximum clade credibility tree from
Wahlberg et al. (2009) \cite{wahlberg2009}. MEDUSA fits alternative birth-death
likelihood models to a phylogenetic tree in order to estimate changes in
net diversification rates along branches. MEDUSA estimates likelihood
and AIC scores for the simplest birth-death model, with two parameters,
the rates \texttt{r}: net diversification \texttt{=} speciation
(\texttt{b}) \texttt{-} extinction (\texttt{d}) and (\(\varepsilon\):
relative extinction \texttt{= d/b}.
The maximum-likelihood values are recalculated when each branch in the tree
is selected to contain an additional breakpoint (additional parameters
\texttt{b} and \texttt{d}) in net diversification rate. 
The process is repeated by fitting additional breaking points to the tree.
The fit of  alternative models is compared using AIC
scores and stop when the improvement of the incrementally more complex models 
is smaller than a cutoff value which
depends of the number of terminals.
MEDUSA finds the likelihood of the models after taking into
account branch lengths and number of species per lineage \cite{alfaro2009}. Most
studies using MEDUSA only run the method on a single tree, usually the
maximum clade credibility tree, which makes the assumption that this
tree is correct. We wanted to study the effects of phylogenetic
uncertainty on estimation of net diversification rate shifts and
therefore selected 1000 random genus-level trees from the posterior
distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
(MultiMEDUSA, S9 File). We calculated a new MCC tree from the
selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
trees and summarized the estimated changes in net diversification rates
for nodes across all trees. Patterns of change in net diversification
rates are considered significant if they are found at the same node in
at least 90\% of the trees. We also expected to find similar values of
net diversification (\texttt{r = b - d}) and relative extinction rate
(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
nodes where changes in diversification tempo occurs. We used the AICc
threshold of 7.8 units, as estimated by MEDUSA, as the limit for a
significantly better fit to select among increasingly complex
alternative models.

@@

As the MEDUSA and MultiMEDUSA approaches estimated an increase in net
diversification in the clade Ithomiini, we tested whether this pattern
can be explained by an increase in the birth-rate due to hostplant use
and performed analyses using the BiSSE model \cite{maddison2007} 
as implemented in the R package
\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
was run for 10000 generations discarding the first 7500 as burnin.
Because most of Nymphalidae butterflies are restricted to
use one plant family as hostplant (232 genera use only one plant family
versus 176 that use more than one; based on our data in S14 Dataset),
the character states can be coded as presence/absence, for which the
BiSSE analysis is better suited. BiSSE was designed to test whether a
binary character state has had any effect on increased net
diversification rate for a clade \cite{maddison2007}. We used our compiled data
of hostplant use to produce a binary dataset for the character ``feeding on
the plant family Solanaceae'' with two states (absence = 0; presence =
1) (S16 Dataset). The family Solanaceae include the hostplants of the
diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
We analyzed the data using BiSSE employing the Markov Chain
Monte Carlo algorithm on the maximum clade credibility tree, taking into
account missing taxa by using the parameter ``sampling factor''
(\texttt{sampling.f}) in \texttt{diversitree}. 
We also run a BiSSE analysis
to test whether the trait ``feeding on Apocynaceae'' had any effect on
net diversification rates found for Nymphalidae lineages feeding on
Apocynaceae.
We also used likelihood ratio tests to evaluate the influence of
Solanaceae-feeding on diversification rates by comparing a constrained versus
a relaxed model of diversification.
The constrained model assumes no effect of hostplant use on diversification by
enforcing equal speciation rates across both character states. In the relaxed
model, all speciation and extinction parameters are estimated for each
character state.
The analyses were run across a sample of 250
trees from the selected 1000 random trees from the posterior
distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
that these species are able to feed on members of this plant family. We tested
whether coding these two genera as not feeding on Solanaceae affected
our results. We expected that coding these two genera as either absence
or presence would not have a significant effect on our results. It has
been shown that BiSSE performs poorly under certain conditions \cite{davis2013}.
However, our data has adequate number of taxa under analysis (more than
300 tips), adequate speciation bias (between 1.5x and 2.0x), character
state bias (around 8x) and extinction bias (around 4x) for the analysis
of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
results \cite{davis2013}.

@@

We used MEDUSA to find out whether taking into account the phylogenetic
signal from the random sample of 1000 trees from the posterior
distribution can return similar estimates of diversification to the
values obtained from the MCC tree. We ran MultiMEDUSA on the random
sample of 1000 trees (S9 File) from the posterior distribution and
compared the results with a MEDUSA analysis on the MCC tree derived from
this sample (S19 Supporting Information).

We found that the analysis by MultiMEDUSA on the 1000 trees estimated
lower median net diversification rates for the diversification shifts
found by MEDUSA on the MCC tree derived from the random sample of trees
(Table~\ref{table2}). Although the diversification pattern found by MEDUSA and
MultiMEDUSA was the same, the latter consistently estimated lower rates.
Furthermore, the shifts recovered with low net diversification rate on
the MCC were recovered with negative net diversification rate by
MultiMEDUSA. The background diversification and all shifts found by
MEDUSA on the 1000 trees are provided as an R object in S12 File.

@@

MultiMEDUSA provided mean and standard deviation statistics for the
diversification values on the shifts on the 1000 trees 
(S6 Fig. and S23 Text), and found that the estimated net diversification rate
values had great variation across the posterior distribution of trees. A
boxplot of the net diversification rate values estimated for the clades
that appear in the MCC tree shows that some shifts are estimated as
increased or slowed diversification pace depending on the tree used for
analysis (Fig.~\ref{fig2}). This variation is especially wide for the clade
formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
because MEDUSA estimated diversification values from six times the
background net diversification rate (\texttt{r = 0.5234}) to almost zero
(\texttt{r = 1.22e-01}). The rates for \emph{Taenaris} were between 0.14
and 0.44 (mean value 0.25). Similar degrees of variation were found in
the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
(Fig.~\ref{fig2}). The net diversification rates estimates for the clades
(Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
Limenitidinae + Heliconiinae and Pseudergolinae are relatively
consistent across the 1000 trees (Fig.~\ref{fig2}).

@@

The MEDUSA analyses, taking into account phylogenetic uncertainty,
estimated a net diversification rate increase in part of the clade
Ithomiini across more than 95\% of the trees. Our BiSSE analysis found a
positive effect of the character state ``feeding on Solanaceae'' on the
net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+ Napeogenina + Dircennina + Godyridina) (Fig.~\ref{fig4}). The estimated mean
net diversification rate for taxa that do not feed on Solanaceae was
\texttt{r = 0.11} while the net diversification rate for the Solanaceae
feeders was \texttt{r = 0.16} (see S1 Fig. for a boxplot of speciation
and extinction values for the 95\% credibility intervals). The same
analysis considering \emph{Vanessa} and \emph{Hypanartia} as
non-Solanaceae feeders due to dubious records produced the same pattern
and net diversification rates (S2 Fig.). Therefore, the rest of the
analyses were performed assuming these two genera as Solanaceae feeders.
We constrained the BiSSE likelihood model to force equal rates of
speciation for both character states in order to test whether the model
of different speciation rates is a significantly better explanation for
the data. A likelihood ratio test found that the model for increased net
diversification rate for nymphalids feeding on Solanaceae is a
significantly better explanation than this character state having no
effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\))
(Table~\ref{table3};
character states available in S16 Dataset, code in S20 Supporting Information,
and mcmc run in S21 Supporting Information). We combined the post-burnin mcmc generations
from running BiSSE on 250 trees from the random sample of 1000 trees and
found the same pattern as the BiSSE analysis on the maximum clade
credibility tree (combined mcmc run in S22 Supporting Information; profiles
plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
S4 Fig.). A BiSSE analysis to test whether the trait ``feeding on
Apocynaceae'' had any effect on increased net diversification rates
found similar speciation rates for lineages feeding on Apocynaceae and
other plants (S5 Fig.).

@@

\subsection*{S6 Fig.}
#\subsection*{S17 Fig.}
\label{S6_Figure}
{\bf Probability of nodes with
estimated rates from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution}.

\subsection*{S7 Fig.}
#\subsection*{S23 Fig.}
\label{S7_Figure}
{\bf Summary of the MultiMEDUSA
analysis on the 1000 trees from the posterior distribution}.

\subsection*{S8 Fig.}
#\subsection*{S25 Fig.}
\label{S8_Figure}
{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.

\subsection*{S9 File}
#\subsection*{S6 File}
\label{S9_File}
{\bf 1000 random trees from \cite{wahlberg2009}}.

\subsection*{S10 File}
\label{S10_File}
{\bf R script to run MEDUSA on the MCC Nymphalidae tree from
\cite{wahlberg2009}}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

\subsection*{S11 File}
\label{S11_File}
{\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
\cite{wahlberg2009}}.

\subsection*{S12 File}
#\subsection*{S15 File}
\label{S12_File}
{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
the posterior distribution}.

\subsection*{S13 Dataset}
#\subsection*{S7 Dataset}
\label{S13_Dataset}
{\bf Species richness for lineages in Nymphalidae}.

\subsection*{S14 Dataset}
#\subsection*{S8 Dataset}
\label{S14_Dataset}
{\bf Hostplants of Nymphalidae
butterflies recorded from the literature}.

\subsection*{S15 Dataset}
#\subsection*{S9 Dataset}
\label{S15_Dataset}
{\bf References for hostplants data}.

\subsection*{S16 Dataset}
#\subsection*{S13 Dataset}
\label{S16_Dataset}
{\bf Data matrix with character states for hosplant use}.

\subsection*{S17 Dataset}
#\subsection*{S21 Dataset}
\label{S17_Dataset}
{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}

\subsection*{S18 Supporting Information}
#\subsection*{S12 Supporting Information}
\label{S18_Supporting_Information}
{\bf R script to run a MEDUSA
analysis on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

\subsection*{S19 Supporting Information}
#\subsection*{S14 Supporting Information}
\label{S19_Supporting_Information}
{\bf MCC tree from the 1000 random
trees selected from the posterior distribution}.

\subsection*{S20 Supporting Information}
#\subsection*{S18 Supporting Information}
\label{S20_Supporting_Information}
{\bf R script for running the BiSSE analysis}.

\subsection*{S21 Supporting Information}
#\subsection*{S19 Supporting Information}
\label{S21_Supporting_Information}
{\bf Raw results for the BiSSE analysis}.

\subsection*{S22 Supporting Information}
#\subsection*{S20 Supporting Information}
\label{S22_Supporting_Information}
{\bf Raw results for the combined BiSSE analysis}.

\subsection*{S23 Text}
#\subsection*{S16 Text}
\label{S23_Text}
{\bf Summary results from a
MultiMEDUSA run on the 1000 random trees selected from the posterior
distribution}.

\subsection*{S24 Text}
#\subsection*{S22 Text}
\label{S24_Text}
{\bf Results of running MEDUSA on the MCC tree}.

\subsection*{S25 Text}
\label{S25_Text}
{\bf Summary results from a MEDUSA
run on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

>>>>

For analyses, we used the phylogenetic trees from the study of Wahlberg
et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
employed the maximum clade credibility tree (MCC tree) (Fig.~\ref{fig1}) as well
as a random sample of 1000 trees from their BEAST run after burnin
\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
used a burnin of 25 million generations and took a random sample of 1000
trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S1 File)
in order to account for phylogenetic uncertainty when performing
the diversification analyses.

@@

Species richness data for Nymphalidae genera were compiled from several
sources including the specialist-curated lists 
\cite{tolweb2007, lamas2004} and curated lists of the
Global Butterfly Names project \cite{gbnp}.
We assigned the species numbers of genera not included in the phylogeny
to the closest related genus that was included in Wahlberg et al. (2009)
\cite{wahlberg2009} study according to available phylogenetic studies 
\cite{matos2013, brower2010, kodandaramaiah2010, kodandaramaiah2010a,
ortiz2013, desilva2010, freitas2004, pena2006, penz1999, silva2008,
pena2011, pena2010},
taxonomical classification and morphological resemblance when no
phylogenies were available (S1 Dataset).

Hostplant data for Nymphalidae species were compiled from several
sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
\cite{dyer2002, beccaloni2008, janzen2009} and others 
(S2 and S3 Dataset) for a
total of 6586
hostplant records, including 428 Nymphalidae genera and 143 plant
families and 1070 plant genera. It was not possible to find any
hostplant data for 35 butterfly genera.

@@

We used the statistical software R version 3.0.1 \cite{r2013} in combination
with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA
\cite{alfaro2009} and
\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
perform the analyses (included as S2 and S3 File). All
analyses were run on the 1000 random trees from Wahlberg et al. (2009)
\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
the MCC tree \cite{wahlberg2009}.

@@

Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
version 093.4.33 \cite{alfaro2009} on the maximum clade credibility tree from
Wahlberg et al. (2009) \cite{wahlberg2009}. MEDUSA fits alternative birth-death
likelihood models to a phylogenetic tree in order to estimate changes in
net diversification rates along branches. MEDUSA estimates likelihood
and AIC scores for the simplest birth-death model, with two parameters,
the rates \texttt{r}: net diversification \texttt{=} speciation
(\texttt{b}) \texttt{-} extinction (\texttt{d}) and (\(\varepsilon\):
relative extinction \texttt{= d/b}.
The maximum-likelihood values are recalculated when each branch in the tree
is selected to contain an additional breakpoint (additional parameters
\texttt{b} and \texttt{d}) in net diversification rate. 
The process is repeated by fitting additional breaking points to the tree.
The fit of  alternative models is compared using AIC
scores and stop when the improvement of the incrementally more complex models 
is smaller than a cutoff value which
depends of the number of terminals.
MEDUSA finds the likelihood of the models after taking into
account branch lengths and number of species per lineage \cite{alfaro2009}. Most
studies using MEDUSA only run the method on a single tree, usually the
maximum clade credibility tree, which makes the assumption that this
tree is correct. We wanted to study the effects of phylogenetic
uncertainty on estimation of net diversification rate shifts and
therefore selected 1000 random genus-level trees from the posterior
distribution of the Bayesian run from Wahlberg et al., 2009 \cite{wahlberg2009}
(MultiMEDUSA, S1 File). We calculated a new MCC tree from the
selection of trees, ran MEDUSA and MultiMEDUSA on the selection of 1000
trees and summarized the estimated changes in net diversification rates
for nodes across all trees. Patterns of change in net diversification
rates are considered significant if they are found at the same node in
at least 90\% of the trees. We also expected to find similar values of
net diversification (\texttt{r = b - d}) and relative extinction rate
(\(\varepsilon\) \texttt{= d/b}) values across the 1000 trees for the
nodes where changes in diversification tempo occurs. We used the AICc
threshold of 7.8 units, as estimated by MEDUSA, as the limit for a
significantly better fit to select among increasingly complex
alternative models.

@@

As the MEDUSA and MultiMEDUSA approaches estimated an increase in net
diversification in the clade Ithomiini, we tested whether this pattern
can be explained by an increase in the birth-rate due to hostplant use
and performed analyses using the BiSSE model \cite{maddison2007} 
as implemented in the R package
\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
was run for 10000 generations discarding the first 7500 as burnin.
Because most of Nymphalidae butterflies are restricted to
use one plant family as hostplant (232 genera use only one plant family
versus 176 that use more than one; based on our data in S2 Dataset),
the character states can be coded as presence/absence, for which the
BiSSE analysis is better suited. BiSSE was designed to test whether a
binary character state has had any effect on increased net
diversification rate for a clade \cite{maddison2007}. We used our compiled data
of hostplant use to produce a binary dataset for the character ``feeding on
the plant family Solanaceae'' with two states (absence = 0; presence =
1) (S4 Dataset). The family Solanaceae include the hostplants of the
diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
We analyzed the data using BiSSE employing the Markov Chain
Monte Carlo algorithm on the maximum clade credibility tree, taking into
account missing taxa by using the parameter ``sampling factor''
(\texttt{sampling.f}) in \texttt{diversitree}. 
We also run a BiSSE analysis
to test whether the trait ``feeding on Apocynaceae'' had any effect on
net diversification rates found for Nymphalidae lineages feeding on
Apocynaceae.
We also used likelihood ratio tests to evaluate the influence of
Solanaceae-feeding on diversification rates by comparing a constrained versus
a relaxed model of diversification.
The constrained model assumes no effect of hostplant use on diversification by
enforcing equal speciation rates across both character states. In the relaxed
model, all speciation and extinction parameters are estimated for each
character state.
The analyses were run across a sample of 250
trees from the selected 1000 random trees from the posterior
distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
that these species are able to feed on members of this plant family. We tested
whether coding these two genera as not feeding on Solanaceae affected
our results. We expected that coding these two genera as either absence
or presence would not have a significant effect on our results. It has
been shown that BiSSE performs poorly under certain conditions \cite{davis2013}.
However, our data has adequate number of taxa under analysis (more than
300 tips), adequate speciation bias (between 1.5x and 2.0x), character
state bias (around 8x) and extinction bias (around 4x) for the analysis
of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
results \cite{davis2013}.

@@

We used MEDUSA to find out whether taking into account the phylogenetic
signal from the random sample of 1000 trees from the posterior
distribution can return similar estimates of diversification to the
values obtained from the MCC tree. We ran MultiMEDUSA on the random
sample of 1000 trees (S1 File) from the posterior distribution and
compared the results with a MEDUSA analysis on the MCC tree derived from
this sample (S2 Supporting Information).

We found that the analysis by MultiMEDUSA on the 1000 trees estimated
lower median net diversification rates for the diversification shifts
found by MEDUSA on the MCC tree derived from the random sample of trees
(Table~\ref{table2}). Although the diversification pattern found by MEDUSA and
MultiMEDUSA was the same, the latter consistently estimated lower rates.
Furthermore, the shifts recovered with low net diversification rate on
the MCC were recovered with negative net diversification rate by
MultiMEDUSA. The background diversification and all shifts found by
MEDUSA on the 1000 trees are provided as an R object in S4 File.

@@

MultiMEDUSA provided mean and standard deviation statistics for the
diversification values on the shifts on the 1000 trees 
(S6 Fig. and S1 Text), and found that the estimated net diversification rate
values had great variation across the posterior distribution of trees. A
boxplot of the net diversification rate values estimated for the clades
that appear in the MCC tree shows that some shifts are estimated as
increased or slowed diversification pace depending on the tree used for
analysis (Fig.~\ref{fig2}). This variation is especially wide for the clade
formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
because MEDUSA estimated diversification values from six times the
background net diversification rate (\texttt{r = 0.5234}) to almost zero
(\texttt{r = 1.22e-01}). The rates for \emph{Taenaris} were between 0.14
and 0.44 (mean value 0.25). Similar degrees of variation were found in
the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
(Fig.~\ref{fig2}). The net diversification rates estimates for the clades
(Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
Limenitidinae + Heliconiinae and Pseudergolinae are relatively
consistent across the 1000 trees (Fig.~\ref{fig2}).

@@

The MEDUSA analyses, taking into account phylogenetic uncertainty,
estimated a net diversification rate increase in part of the clade
Ithomiini across more than 95\% of the trees. Our BiSSE analysis found a
positive effect of the character state ``feeding on Solanaceae'' on the
net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+ Napeogenina + Dircennina + Godyridina) (Fig.~\ref{fig4}). The estimated mean
net diversification rate for taxa that do not feed on Solanaceae was
\texttt{r = 0.11} while the net diversification rate for the Solanaceae
feeders was \texttt{r = 0.16} (see S1 Fig. for a boxplot of speciation
and extinction values for the 95\% credibility intervals). The same
analysis considering \emph{Vanessa} and \emph{Hypanartia} as
non-Solanaceae feeders due to dubious records produced the same pattern
and net diversification rates (S2 Fig.). Therefore, the rest of the
analyses were performed assuming these two genera as Solanaceae feeders.
We constrained the BiSSE likelihood model to force equal rates of
speciation for both character states in order to test whether the model
of different speciation rates is a significantly better explanation for
the data. A likelihood ratio test found that the model for increased net
diversification rate for nymphalids feeding on Solanaceae is a
significantly better explanation than this character state having no
effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\))
(Table~\ref{table3};
character states available in S4 Dataset, code in S3 Supporting Information,
and mcmc run in S4 Supporting Information). We combined the post-burnin mcmc generations
from running BiSSE on 250 trees from the random sample of 1000 trees and
found the same pattern as the BiSSE analysis on the maximum clade
credibility tree (combined mcmc run in S5 Supporting Information; profiles
plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
S4 Fig.). A BiSSE analysis to test whether the trait ``feeding on
Apocynaceae'' had any effect on increased net diversification rates
found similar speciation rates for lineages feeding on Apocynaceae and
other plants (S5 Fig.).

@@

\subsection*{S6 Fig.}
\label{S6_Figure}
{\bf Probability of nodes with
estimated rates from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution}.

\subsection*{S7 Fig.}
\label{S7_Figure}
{\bf Summary of the MultiMEDUSA
analysis on the 1000 trees from the posterior distribution}.

\subsection*{S8 Fig.}
\label{S8_Figure}
{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.

\subsection*{S1 File}
\label{S1_File}
{\bf 1000 random trees from \cite{wahlberg2009}}.

\subsection*{S2 File}
\label{S2_File}
{\bf R script to run MEDUSA on the MCC Nymphalidae tree from
\cite{wahlberg2009}}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

\subsection*{S3 File}
\label{S3_File}
{\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
\cite{wahlberg2009}}.

\subsection*{S4 File}
\label{S4_File}
{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
the posterior distribution}.

\subsection*{S1 Dataset}
\label{S1_Dataset}
{\bf Species richness for lineages in Nymphalidae}.

\subsection*{S2 Dataset}
\label{S2_Dataset}
{\bf Hostplants of Nymphalidae
butterflies recorded from the literature}.

\subsection*{S3 Dataset}
\label{S3_Dataset}
{\bf References for hostplants data}.

\subsection*{S4 Dataset}
\label{S4_Dataset}
{\bf Data matrix with character states for hosplant use}.

\subsection*{S5 Dataset}
\label{S5_Dataset}
{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}

\subsection*{S1 Supporting Information}
\label{S1_Supporting_Information}
{\bf R script to run a MEDUSA
analysis on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

\subsection*{S2 Supporting Information}
\label{S2_Supporting_Information}
{\bf MCC tree from the 1000 random
trees selected from the posterior distribution}.

\subsection*{S3 Supporting Information}
\label{S3_Supporting_Information}
{\bf R script for running the BiSSE analysis}.

\subsection*{S4 Supporting Information}
\label{S4_Supporting_Information}
{\bf Raw results for the BiSSE analysis}.

\subsection*{S5 Supporting Information}
\label{S5_Supporting_Information}
{\bf Raw results for the combined BiSSE analysis}.

\subsection*{S1 Text}
\label{S1_Text}
{\bf Summary results from a
MultiMEDUSA run on the 1000 random trees selected from the posterior
distribution}.

\subsection*{S2 Text}
\label{S2_Text}
{\bf Results of running MEDUSA on the MCC tree}.

\subsection*{S3 Text}
\label{S3_Text}
{\bf Summary results from a MEDUSA
run on the MCC tree from the 1000 random trees selected from the
posterior distribution}.
